---
output:
  md_document:
    variant: gfm
html_preview: false
---


```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
library("ypages")

Biocpkg <- function (pkg) {
    sprintf("[%s](http://bioconductor.org/packages/%s)", pkg, pkg)
}
```

#  GOSemSim: GO semantic similarity measurement


[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#since)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![platform](http://www.bioconductor.org/shields/availability/devel/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#archives)
[![codecov](https://codecov.io/gh/GuangchuangYu/GOSemSim/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/GOSemSim/)


`r badge_bioc_release("GOSemSim", "green")`
`r badge_devel("guangchuangyu/GOSemSim", "green")`
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/GOSemSim/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/GOSemSim)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/GOSemSim/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/GOSemSim)


`r badge_bioc_download("GOSemSim", "total", "blue")`
`r badge_bioc_download("GOSemSim", "month", "blue")`
`r badge_bioc_download_rank("GOSemSim")`

```{r comment="", echo=FALSE, results='asis'}
cat(packageDescription('GOSemSim')$Description)
```


[![Twitter](https://img.shields.io/twitter/url/http/shields.io.svg?style=social&logo=twitter)](https://twitter.com/intent/tweet?hashtags=GOSemSim&url=http://bioinformatics.oxfordjournals.org/content/26/7/976&screen_name=guangchuangyu)
[![saythanks](https://img.shields.io/badge/say-thanks-ff69b4.svg)](https://saythanks.io/to/GuangchuangYu)
`r badger::badge_custom("follow me on", "WeChat", "green", "https://guangchuangyu.github.io/blog_images/biobabble.jpg")`



## :writing_hand: Authors

Guangchuang YU <https://guangchuangyu.github.io>

School of Basic Medical Sciences, Southern Medical University


If you use `r Biocpkg('GOSemSim')` in published research, please cite:


+ **Yu G**. [Gene Ontology Semantic Similarity Analysis Using GOSemSim](http://dx.doi.org/10.1007/978-1-0716-0301-7_11). In: Kidder B. (eds) Stem Cell Transcriptional Networks. **_Methods in Molecular Biology_**, 2020, 2117:207-215. Humana, New York, NY. 
+ **Yu G**^#^, Li F^#^, Qin Y, Bo X^\*^, Wu Y and Wang S^\*^. [GOSemSim: an R package for measuring semantic similarity among GO terms and gene products](http://dx.doi.org/10.1093/bioinformatics/btq064). **_Bioinformatics_**. 2010, 26(7):976-978. 


## :arrow_double_down: Installation

Get the released version from Bioconductor:

```r
## try http:// if https:// URLs are not supported
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
## BiocManager::install("BiocUpgrade") ## you may need this
BiocManager::install("GOSemSim")
```

Or the development version from github:

```r
## install.packages("remotes")
remotes::install_github("GuangchuangYu/GOSemSim")
```



## :sparkling_heart: Contributing

We welcome any contributions! By participating in this project you agree to
abide by the terms outlined in the [Contributor Code of Conduct](CONDUCT.md).

