---
output:
  md_document:
    variant: markdown_github
html_preview: false
---

#  GOSemSim: GO semantic similarity measurement

[![platform](http://www.bioconductor.org/shields/availability/devel/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/GOSemSim.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/GOSemSim/)
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/GOSemSim/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/GOSemSim)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/GOSemSim/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/GOSemSim)
[![install with bioconda](https://img.shields.io/badge/install%20with-bioconda-green.svg?style=flat)](http://bioconda.github.io/recipes/bioconductor-gosemsim/README.html)

[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![codecov](https://codecov.io/gh/GuangchuangYu/GOSemSim/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/GOSemSim/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/GOSemSim/commits/master)
[![commit](http://www.bioconductor.org/shields/commits/bioc/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#svn_source)
[![GitHub forks](https://img.shields.io/github/forks/GuangchuangYu/GOSemSim.svg)](https://github.com/GuangchuangYu/GOSemSim/network) 
[![GitHub stars](https://img.shields.io/github/stars/GuangchuangYu/GOSemSim.svg)](https://github.com/GuangchuangYu/GOSemSim/stargazers) 


[![releaseVersion](https://img.shields.io/badge/release%20version-`r url="https://bioconductor.org/packages/release/bioc/html/GOSemSim.html"; x=readLines(url); gsub("\\D+(\\d+\\.\\d+\\.\\d+)\\D+", "\\1", x[grep("Version", x) +1])`-green.svg?style=flat)](https://bioconductor.org/packages/GOSemSim)
[![develVersion](https://img.shields.io/badge/devel%20version-`r x=readLines("DESCRIPTION"); y=x[grep("^Version", x)]; sub("Version: ", "", y)`-green.svg?style=flat)](https://github.com/GuangchuangYu/GOSemSim)
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#since)
[![post](http://www.bioconductor.org/shields/posts/GOSemSim.svg)](https://support.bioconductor.org/t/GOSemSim/)
[![download](http://www.bioconductor.org/shields/downloads/GOSemSim.svg)](https://bioconductor.org/packages/stats/bioc/GOSemSim/)


The semantic comparisons of Gene Ontology (GO) annotations provide quantitative ways to compute similarities between genes and gene groups, and have became important basis for many bioinformatics analysis approaches. `GOSemSim` is an R package for semantic similarity computation among GO terms, sets of GO terms, gene products and gene clusters. `GOSemSim` implemented five methods proposed by *Resnik*, *Schlicker*, *Jiang*, *Lin* and *Wang* respectively.

[![Twitter](https://img.shields.io/twitter/url/https/github.com/GuangchuangYu/ggtree.svg?style=social)](https://twitter.com/intent/tweet?hashtags=GOSemSim&url=https://guangchuangyu.github.io/GOSemSim)

------------------------------------------------------------------------

Please cite the following article when using `GOSemSim`: 

**Yu G**^†^, Li F^†^, Qin Y, Bo X^\*^, Wu Y and Wang S^\*^. GOSemSim: an R package for measuring semantic similarity among GO terms and gene products. ***Bioinformatics***. 2010, 26(7):976-978. 

[![doi](https://img.shields.io/badge/doi-10.1093/bioinformatics/btq064-green.svg?style=flat)](http://dx.doi.org/10.1093/bioinformatics/btq064) 
[![citation](https://img.shields.io/badge/cited%20by-`r library(scholar); sum(get_article_cite_history('DO5oG40AAAAJ', 'tuHXwOkdijsC')$cites)`-green.svg?style=flat)](https://scholar.google.com.hk/scholar?hl=en&as_sdt=0,5&sciodt=0,5&cites=9484177541993722322,17633835198940746971,18126401808149291947&scipsc=)

------------------------------------------------------------------------

For details, please visit our project website, <https://guangchuangyu.github.io/GOSemSim>.

+ [Documentation](https://guangchuangyu.github.io/GOSemSim/documentation/)
+ [Featured Articles](https://guangchuangyu.github.io/GOSemSim/featuredArticles/)
+ [Feedback](https://guangchuangyu.github.io/GOSemSim/#feedback)

### Citation 

[![citation](https://img.shields.io/badge/cited%20by-`r library(scholar); sum(get_article_cite_history('DO5oG40AAAAJ', 'tuHXwOkdijsC')$cites)`-blue.svg?style=flat)](https://scholar.google.com.hk/scholar?hl=en&as_sdt=0,5&sciodt=0,5&cites=9484177541993722322,17633835198940746971,18126401808149291947&scipsc=)

```{r echo=F, comment=NA}
d <- scholar::get_article_cite_history('DO5oG40AAAAJ', 'tuHXwOkdijsC')
txtplot::txtplot(d$year, d$cites)
```

### Download stats

[![download](http://www.bioconductor.org/shields/downloads/GOSemSim.svg)](https://bioconductor.org/packages/stats/bioc/GOSemSim/)
[![total](https://img.shields.io/badge/downloads-`r sum(dlstats::bioc_stats("GOSemSim")$Nb_of_downloads)`/tatal-blue.svg?style=flat)](https://bioconductor.org/packages/stats/bioc/GOSemSim/)
[![month](https://img.shields.io/badge/downloads-`r x <- dlstats::bioc_stats("GOSemSim"); x$Nb_of_downloads[nrow(x)-1]`/month-blue.svg?style=flat)](https://bioconductor.org/packages/stats/bioc/GOSemSim/)

```{r echo=F, comment=NA}
x <- dlstats::bioc_stats("GOSemSim")
x <- x[-nrow(x), ] # last month data is not complete
txtplot::txtplot(ggtree::Date2decimal(x$end), x$Nb_of_downloads, width=100, height=25)
```



