---
output:
  md_document:
    variant: markdown_github
html_preview: false
---


```{r echo=FALSE, results="hide", message=FALSE}
library("ypages")
library("badger")
library("txtplot")
```

#  GOSemSim: GO semantic similarity measurement

`r badge_bioc_release("GOSemSim", "green")`
`r badge_devel("guangchuangyu/GOSemSim", "green")`
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#since)
`r badge_bioc_download("GOSemSim", "total", "blue")`
`r badge_bioc_download("GOSemSim", "month", "blue")`


[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![codecov](https://codecov.io/gh/GuangchuangYu/GOSemSim/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/GOSemSim/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/GOSemSim/commits/master)
[![commit](http://www.bioconductor.org/shields/commits/bioc/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#svn_source)
[![GitHub forks](https://img.shields.io/github/forks/GuangchuangYu/GOSemSim.svg)](https://github.com/GuangchuangYu/GOSemSim/network)
[![GitHub stars](https://img.shields.io/github/stars/GuangchuangYu/GOSemSim.svg)](https://github.com/GuangchuangYu/GOSemSim/stargazers)


[![platform](http://www.bioconductor.org/shields/availability/devel/GOSemSim.svg)](https://www.bioconductor.org/packages/devel/bioc/html/GOSemSim.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/GOSemSim.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/GOSemSim/)
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/GOSemSim/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/GOSemSim)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/GOSemSim/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/GOSemSim)



The semantic comparisons of Gene Ontology (GO) annotations provide quantitative ways to compute similarities between genes and gene groups, and have became important basis for many bioinformatics analysis approaches. `GOSemSim` is an R package for semantic similarity computation among GO terms, sets of GO terms, gene products and gene clusters. `GOSemSim` implemented five methods proposed by *Resnik*, *Schlicker*, *Jiang*, *Lin* and *Wang* respectively.

[![Twitter](https://img.shields.io/twitter/url/https/github.com/GuangchuangYu/GOSemSim.svg?style=social)](https://twitter.com/intent/tweet?hashtags=GOSemSim&url=http://bioinformatics.oxfordjournals.org/content/26/7/976&screen_name=guangchuangyu)

------------------------------------------------------------------------

Please cite the following article when using `GOSemSim`:

**Yu G**^†^, Li F^†^, Qin Y, Bo X^\*^, Wu Y and Wang S^\*^. GOSemSim: an R package for measuring semantic similarity among GO terms and gene products. ***Bioinformatics***. 2010, 26(7):976-978.


`r badge_doi("10.1093/bioinformatics/btq064", "green")`
`r badge_altmetric("100979", "green")`
`r badge_citation("tuHXwOkdijsC", "9484177541993722322", "green")`
`r badge_custom("ESI", "Highly Cited Paper", "green", "http://apps.webofknowledge.com/InboundService.do?mode=FullRecord&customersID=RID&IsProductCode=Yes&product=WOS&Init=Yes&Func=Frame&DestFail=http%3A%2F%2Fwww.webofknowledge.com&action=retrieve&SrcApp=RID&SrcAuth=RID&SID=Y2CXu6nry8nDQZcUy1w&UT=WOS%3A000276045800023")`

------------------------------------------------------------------------

For details, please visit our project website, <https://guangchuangyu.github.io/GOSemSim>.

+ [Documentation](https://guangchuangyu.github.io/GOSemSim/documentation/)
+ [Featured Articles](https://guangchuangyu.github.io/GOSemSim/featuredArticles/)
+ [Feedback](https://guangchuangyu.github.io/GOSemSim/#feedback)

### Citation

`r badge_citation("tuHXwOkdijsC", "9484177541993722322", "green")`
`r badge_custom("ESI", "Highly Cited Paper", "green", "http://apps.webofknowledge.com/InboundService.do?mode=FullRecord&customersID=RID&IsProductCode=Yes&product=WOS&Init=Yes&Func=Frame&DestFail=http%3A%2F%2Fwww.webofknowledge.com&action=retrieve&SrcApp=RID&SrcAuth=RID&SID=Y2CXu6nry8nDQZcUy1w&UT=WOS%3A000276045800023")`


```{r echo=F, comment=NA}
d <- get_article_cite_history('DO5oG40AAAAJ', 'tuHXwOkdijsC')
txtplot(d$year, d$cites)
```

### Download stats

[![download](http://www.bioconductor.org/shields/downloads/GOSemSim.svg)](https://bioconductor.org/packages/stats/bioc/GOSemSim/)
`r badge_bioc_download("GOSemSim", "total", "blue")`
`r badge_bioc_download("GOSemSim", "month", "blue")`


```{r echo=F, comment=NA}
x <- dlstats::bioc_stats("GOSemSim")
x <- x[-nrow(x), ] # last month data is not complete
txtplot(ggtree::Date2decimal(x$end), x$Nb_of_downloads, width=100, height=25)
```



